### 2017-08-31 paper.tex
:computer: Mechanics of collapse

\usepackage[backend=biber, style=chem-angew, citestyle=authoryear, sorting=nty, natbib=true]{biblatex}
\bibliography{references} %Location of references.bib only for biblatex

@@

>>>>

\usepackage[backend=biber, style=numeric, citestyle=authoryear, sorting=nty, natbib=true]{biblatex}
\bibliography{references} %Location of references.bib only for biblatex

@@

\subsubsection{Turbulence and drag}
Closer observation of the surface of the flow front in the ~\cref{fig:dense-s5-r09} and the loose column cases~\cref{fig:slope_loose_5} cases shows that the number of turbulent vortices formed on the surface of the flow is significantly higher in the dense column cases than in the loose column cases. This can be attributed to the triangular profiles of the flowing mass of the dense column cases, in contrast to the more parabolic profiles observed in the loose granular column cases. This increase in the interaction between the flowing granular mass and the neighbouring fluid results in an increased momentum transfer and thus formation of turbulent vortices.~\Cref{fig:water_entrainment_a08_dense} shows higher density of streamlines at the top surface for the dense granular flow, which indicates increased drag resistance. This in turn reduces the velocity and the final runout distance. Viscous drag effect is found to reduce the acceleration up to about 30\% in underwater landslides~\citep{Watts2000}. In contrast the loose granular flowing mass experiences less drag, thereby reaching higher kinetic energy and longer runout distances. These mechanisms explain the loose granular masses in fluid spreading almost twice as long as the dense granular masses, as also observed in experiments~\citep{Iverson2000, Schaeffer2008}.

\subsection{Settlement phase}
In both dense and loose cases, initially the packing density decreases with time during the runout initiation process ($t = 0$ to $3\tau_c$). As the body starts to slow down, the density then increases. At the end of the flow, both the dense and the loose cases reach a similar packing density, indicating that the initial density effect is erased and the granular mass reaches the critical state~\citep{Schofield1968}. As the granular mass settles, the eddies which developed during the horizontal acceleration phase starts to depart away from the granular surface. Force chains start to reappear at this phase, as individual particles re-establish contacts, and the shear resistance of the granular mass increases.
